$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-09-03 paper/paper.tex
Provide a more realistic example

\emph{In this paper we present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}

@@

\begin{example}[Variable reference]
  \autoref{lst:channels} types a variable reference under a linear algebra.
  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
_ : $\contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
_ = $\suc \constr{0}$

_ : $\contains{\constr{[]} \comma \channel{\unit}{\li}} {\constr{[]} \comma \lio} {\constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo}$
_ = $\constr{0}$
  \end{lstlisting}
  The underscore \_ introduces an anonymous declaration immediately followed by its definition.
  We must show that the variable at position $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $\constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\constr{[]} \comma \lio \comma \lio$.
  The \textsc{VarRef} constructors we must use are completely determined by $\suc \; \constr{0}$ in the type.
  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output, the result is the obligation on line 4.
  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to fulfill this proof obligation automatically.
\end{example}

@@

    \inferrule
        {\stacked{
            chan_i : \contains{\gamma \hspace{1.2em}}{\Gamma \hspace{1.3em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
            cont \hspace{0.6em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.6em}}{\Theta \comma \lz}}}
        {\recv{chan_i}{} cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}

    \inferrule
        {\stacked{
            chan_i : \contains{\gamma}{\Gamma \hspace{0.2em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
            loc_j \hspace{0.8em} : \contains{\gamma}{\Delta}{j}{t \hspace{3.2em}}{x \hspace{0.3em}}{\Xi} \\\\
            cont \hspace{0.6em} : \types{\gamma}{\Xi \hspace{0.3em}}{P\hspace{4.8em}}{\Theta}}}
        {\send{chan_i}{loc_j} cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}

    \inferrule
    {l \hspace{0.3em} : \types{\gamma}{\Gamma \hspace{0.2em}}{P}{\Delta} \\\\
     r : \types{\gamma}{\Delta}{Q}{\Xi}}
    {\comp{l}{r} : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
  \end{mathpar}
  \caption{Leftover typing for a resource-aware type system.}
  \label{fig:types}
  \end{figure}

@@

  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing two processes that send parts of one channel over another.]
$\func{P}$ : $\Process_{\suc \suc \constr{0}}$
$\func{P}$ = $\comp{(\new{} \; (\send{\suc \; \constr{0}} {\constr{0}} \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))}{\recv{\constr{0}}{\recv{\constr{0}}{}} \PO}$

$\func{\gamma}$ : $\PreCtx_{\suc \suc \constr{0}}$
$\func{\gamma}$ = $\constr{[]} \comma \unit \comma \channel{\channel{\unit}{\func{\omega}}}{\li}$

$\func{\vdash P}$ : $\forall \{n\} \to \types{\func{\gamma}}{\constr{[]} \comma \func{\omega} \comma (\suc n \comma \suc n)}{\func{P}}{\constr{[]} \comma \func{\omega} \comma (n \comma n)}$
$\func{\vdash P}$ = $\comp{(\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))} {\recv{\constr{0}}{} \; \recv{\constr{0}}{} \PO}$

$\func{\vdash PP}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\new{} \comp{\func{P}}{\func{P}}}{\constr{[]} \comma \func{\omega}}$
$\func{\vdash PP}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; (\suc \suc \constr{0}) \; (\comp{\vdash P}{\vdash P})$
  \end{lstlisting}

>>>>

\emph{We present the first full mechanisation in Agda of a \picalc{} with linear, graded and shared types, all under a common unified framework.}

@@

\begin{example}[Variable reference]
  \autoref{lst:channels} types a variable reference under a linear algebra.
  \begin{lstlisting}[label=lst:channels,mathescape,numbers=left,caption=Typing variable reference $\suc \constr{0}$ with type $\channel{\unit}{\li}$ and usage $\li$.]
$\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li} \comma \unit} {\constr{[]} \comma \lio \comma \lio} {\suc \constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo \comma \lio}$
$\_ \; = \; \suc \constr{0}$

$\_ \; : \; \contains{\constr{[]} \comma \channel{\unit}{\li}} {\constr{[]} \comma \lio} {\constr{0}} {\channel{\unit}{\li}} {\li} {\constr{[]} \comma \lo}$
$\_ \; = \; \constr{0}$
  \end{lstlisting}
  The underscore \_ introduces an anonymous declaration immediately followed by its definition.
  We must show that the variable at position $\suc \constr{0}$ is of type $\channel{\unit}{\li}$ and has a usage annotation $\li$ in an environment with a typing context $\constr{[]} \comma \channel{\unit}{\li} \comma \unit$ and a usage context $\constr{[]} \comma \lio \comma \lio$.
  The \textsc{VarRef} constructors we must use are completely determined by $\suc \; \constr{0}$ in the type.
  The constructor $\suc$ steps under the outermost variable in the context, preserving its usage annotation $\lio$ from input to output, the result is the obligation on line 4.
  The constructor $\constr{0}$ asserts that the next variable is of type $\channel{\unit}{\li}$, and that the usage annotation $\lio$ can be split such that $\op{\lio}{\li}{\lo}$ --- we use $\func{\cdot-compute^r}$ to fulfill this proof obligation automatically.
\end{example}

@@

    \inferrule
        {\stacked{
            chan_i : \contains{\gamma \hspace{1.0em}}{\Gamma \hspace{1.0em}}{i}{\channel{t}{x}}{\li}{\Xi} \\\\
            cont \hspace{0.5em} : \types{\gamma \comma t}{\Xi \comma x}{P \hspace{4.2em}}{\Theta \comma \lz}}}
        {\recv{chan_i}{} cont : \types{\gamma}{\Gamma}{\recv{i}{}P}{\Theta}}

    \inferrule
        {\stacked{
            chan_i : \contains{\gamma}{\Gamma \hspace{0.1em}}{i}{\channel{t}{x}}{\lo}{\Delta} \\\\
            loc_j \hspace{0.7em} : \contains{\gamma}{\Delta}{j}{t \hspace{2.8em}}{x \hspace{0.2em}}{\Xi} \\\\
            cont \hspace{0.5em} : \types{\gamma}{\Xi}{\hspace{0.4em}P\hspace{4.0em}}{\Theta}}}
        {\send{chan_i}{loc_j} cont : \types{\gamma}{\Gamma}{\send{i}{j}P}{\Theta}}

    \inferrule
    {l : \types{\gamma}{\Gamma}{P\hspace{0.3em}}{\Delta} \\\\
     r : \types{\gamma}{\Delta}{Q}{\Xi}}
    {\comp{l}{r} : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
  \end{mathpar}
  \caption{Leftover typing for a resource-aware type system.}
  \label{fig:types}
  \end{figure}

@@

  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing two processes that send parts of one channel over another.]
  $\func{sync} \; : \; \forall \{n\} \to \Var_n \to \Var_n \to \Var_n \to \Process_n$
  $\func{sync} \; i0 \; i1 \; o \; =$
    $\recv{i0}{} \recv{(\suc \; i1)}{} \send{(\suc \suc \; o)}{\suc \constr{0}} \send{(\suc \suc \; o)}{\constr{0}} \PO$

  $\func{send} \; : \; \forall \{n\} \to \Var_n \to \Process_n$
  $\func{send} \; c \; = \new{} (\send{\suc c}{\constr{0}} \PO)$

  $\func{recv} \; : \; \forall \{n\} \to \Var_n \to \Process_n$
  $\func{recv} \; c \; = \recv{c}{} (\recv{\suc c} \PO)$

  $\func{example} \; : \; \Process_{\constr{0}}$
  $\func{example} \; = \new{} ( \func{send} \; \constr{0}$
           $\comp{}{ \new{} ( \func{send} \; \constr{0} }$
             $\comp{}{ \new{} ( \func{recv} \; \constr{0} }$
               $\comp{}{ \func{sync} \; (\suc \suc \constr{0}) \; (\suc \constr{0}) \; \constr{0}))) }$

  $\func{\_} \; : \; \types{\constr{[]}}{\constr{[]}}{\func{example}}{\constr{[]}}$
  $\func{\_} = \new{} \; \channel{\unit}{\omega} \; \lz \; 1$
    $( (\send{\suc \constr{0}}{\constr{0}} \PO)$
    $\comp{}{ \new{} \; \channel{\unit}{\omega} \; \lz \; 1 }$
      $( (\send{\suc \constr{0}}{\constr{0}} \PO)$
      $\comp{}{ \new{} \; \channel{\unit}{\omega} \; \lz \; 2 }$
        $( (\recv{\constr{0}}{} (\recv{\suc \constr{0}}{} \PO))$
        $\comp{}{( ( \recv{\suc (\suc \constr{0}))}{}}$
          $( \recv{\suc (\suc \constr{0}))}{}$
          $( \send{\suc (\suc \constr{0}))}{\suc \constr{0}}$
          $( \send{\suc (\suc \constr{0}))}{\constr{0}} \PO))))$
  \end{lstlisting}
